---
title: "MaggotLabExMar2020"
author: "JReceveur"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: hide
---
  
  ```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=14, fig.height=10)
knitr::opts_chunk$set(fig.align="center")
```

``` {r import, message=FALSE, warning=FALSE,echo=TRUE}

#Parse Silva function
parse_taxonomy_silva_128 <- function(char.vec){
  # Use default to assign names to elements in case problem with greengenes prefix
  char.vec = parse_taxonomy_default(char.vec)
  # Check for unassigned taxa
  if (char.vec["Rank1"] == "Unassigned") {
    char.vec <- c(Rank1="D_0__Unassigned", Rank2="D_1__Unassigned", Rank3="D_2__Unassigned", Rank4="D_3__Unassigned",
                  Rank5="D_4__Unassigned", Rank6="D_5__Unassigned", Rank7="D_6__Unassigned")
  }
  # Define the meaning of each prefix according to SILVA taxonomy
  Tranks = c(D_0="Kingdom", D_1="Phylum", D_2="Class", D_3="Order", D_4="Family", D_5="Genus", D_6="Species")
  # Check for prefix using regexp, warn if there were none. trim indices, ti
  ti = grep("[[:alpha:]]\\_[[:digit:]]{1}\\_\\_", char.vec)
  if( length(ti) == 0L ){
    warning(
      "No silva prefixes were found. \n",
      "Consider using parse_taxonomy_delfault() instead if true for all OTUs. \n",
      "Dummy ranks may be included among taxonomic ranks now."
    )
    # Will want to return without further modifying char.vec
    taxvec = char.vec
    # Replace names of taxvec according to prefix, if any present...
  } else {
    # Format character vectors for Ambiguous taxa
    if( length(ti) < 7 ){
      for (key in names(char.vec)) {
        if ( char.vec[key] == "Ambiguous_taxa" ) {
          tax_no <- (as.numeric(substr(key, 5, 5)) - 1)
          char.vec[key] = sprintf("D_%s__Ambiguous_taxa", tax_no)
        }
      }
      # Reset the trimmed indicies if Ambiguous taxa
      ti = grep("[[:alpha:]]\\_[[:digit:]]{1}\\_\\_", char.vec)
    }
    # Remove prefix using sub-"" regexp, call result taxvec
    taxvec = gsub("[[:alpha:]]\\_[[:digit:]]{1}\\_\\_", "", char.vec)
    # Define the ranks that will be replaced
    repranks = Tranks[substr(char.vec[ti], 1, 3)]
    # Replace, being sure to avoid prefixes notK present in Tranks
    names(taxvec)[ti[!is.na(repranks)]] = repranks[!is.na(repranks)]
  }
  return(taxvec)
}

library(vegan)
library(MASS)
library(ggplot2)
library(plyr)
library(dplyr)
library(magrittr)
library(scales)
library(grid)
library(reshape2)
library(phyloseq)
library(randomForest)
library(knitr)
library(ape)
library(ggpubr)
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#000000","#CC79A7")
theme_set(theme_bw(base_size = 18)+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()))
biom=import_biom("LabMaggotMar2020.biom",parseFunction= parse_taxonomy_silva_128)
metadata=read.table("MaggotExpMetadata.txt",header = TRUE)
tree=read_tree("MaggotTreeMar2020.nwk")
sampdat=sample_data(metadata)
sample_names(sampdat)=metadata$SampleID
physeq=merge_phyloseq(biom,sampdat,tree)


physeq
###### rarefaction 4K

```

#Alpha Diversity
###Observed Species

```{r}
plot_richness(physeq, x="Treatment",color="Treatment", measures=c("Observed"))+geom_boxplot(aes(x=Treatment, y=value, color=Treatment), alpha=0.05)+ylab("Observed Species")

```

##Simpson Diversity
```{r}
plot_richness(physeq, x="Treatment",color="Treatment", measures=c("Shannon"),)+geom_boxplot(aes(x=Treatment, y=value, color=Treatment), alpha=0.05)+ylab("Shannon Diversity")
plot_richness(physeq, x="Treatment",color="Treatment", measures=c("Shannon"),)+geom_boxplot(aes(x=Treatment, y=value, color=Treatment), alpha=0.05)+ylab("Shannon Diversity")+facet_wrap(~Type)
plot_richness(physeq, x="Treatment",color="Treatment", measures=c("Simpson"),)+geom_boxplot(aes(x=Treatment, y=value, color=Treatment), alpha=0.05)+ylab("Simpson Diversity")+facet_wrap(~SamplingTime~Treatment)

```


#Taxa Plots

```{r filteringForFamily2}


GPr  = transform_sample_counts(physeq, function(x) x / sum(x) ) #transform samples based on relative abundance
GPrPhylum=tax_glom(GPr, "Phylum")
PhylumLevel = filter_taxa(GPrPhylum, function(x) mean(x) > 1e-2, TRUE) #filter out any taxa lower tha 0.1%
GPrFamily=tax_glom(GPr,"Family")
FamilyLevel = filter_taxa(GPrFamily, function(x) mean(x) > 2e-2, TRUE) #filter out any taxa lower tha 1%
GPrGenus=tax_glom(GPr,"Genus")
GenusLevel = filter_taxa(GPrGenus, function(x) mean(x) > 2e-2, TRUE) #filter out any taxa lower tha 1%

```

##Phylum Level



```{r PhylumRA, warning=FALSE}
df <- psmelt(PhylumLevel)
df$Abundance=df$Abundance*100
Trtdata <- ddply(df, c("Phylum", "Treatment","Type"), summarise,
                 N    = length(Abundance),
                 mean = mean(Abundance),
                 sd   = sd(Abundance),
                 se   = sd / sqrt(N)
)
```

```{r TreatmentPlot5}
#labels <- c(Ceca = "Ceca", Cloacal_Swab = "Cloacal Swab",Large_Intestine_Tissue="Large Intestine Tissue", Small_Intestine_Contents="Small Intestine Contents",Headgut_Swab="Headgut Swab",Small_Intestine_Tissue="Small Intestine Tissue")
cdataplot=ggplot(Trtdata, aes(x=Treatment,y=mean))+geom_bar(aes(fill = Phylum),colour="black", stat="identity")+xlab("Treatment")+ylab("Relative Abundance (> 1%)")+theme(axis.title.x=element_blank())+ theme(axis.text.x = element_text(angle = 45, hjust = 1))+ scale_fill_manual(values=cbPalette)+facet_wrap(~Type)
cdataplot#+  scale_x_discrete(labels=c("0 hrs", "24 hrs", "48 hrs","72 hrs"))
#position = "fill"#+geom_errorbar(aes(ymin=mean-se,ymax=mean+se),color="black",width=1)+geom_line(size=1.5,linetype="dashed")+geom_point(size=6)+ylab(
```

```{r}
df <- psmelt(PhylumLevel)
df$Abundance=df$Abundance*100
Trtdata <- ddply(df, c("Phylum", "Treatment"), summarise,
                 N    = length(Abundance),
                 mean = mean(Abundance),
                 sd   = sd(Abundance),
                 se   = sd / sqrt(N)
)
cdataplot=ggplot(Trtdata, aes(x=Treatment,y=mean))+geom_bar(aes(fill = Treatment),colour="black", stat="identity")+ facet_grid(~Phylum)+xlab("Decomp Stage")+ylab("Relative Abundance (> 2%)") + theme(axis.text.x = element_text(angle = 45, hjust = 1))+theme(axis.title.x=element_blank())+facet_wrap(~Phylum)+geom_errorbar(aes(ymin=mean-se,ymax=mean+se))+scale_colour_manual(values=cbPalette)+scale_fill_manual(values=cbPalette)
#scale_x_discrete(labels=c("0 hrs", "24 hrs", "48 hrs","72 hrs"))+ theme(axis.text.x = element_text(angle = 45, hjust = 1))#+ scale_fill_manual(values=cbPalette)
cdataplot
Means=compare_means(Abundance ~ Treatment, data = df, 
                    group.by = "Phylum", p.adjust.method = "fdr")
#head(Means)
keeps <- c("Phylum","group1","group2","p.adj","method","p.signif")
keeps=Means[keeps]
#keeps


test3 <- list('Phylum'= keeps$Phylum,'group1'=keeps$group1,'group2'= keeps$group2,'p.adj'=keeps$p.adj,p.signif=keeps$p.signif,'Method'=keeps$method)
test3= as.data.frame(test3)
#test3
FilteredResults<-test3[!(test3$p.adj>0.1),]            
FilteredResults
```


##Family Level Relative Abundance


```{r Summarizing2,warning=FALSE}
df <- psmelt(FamilyLevel)
df$Abundance=df$Abundance*100
Trtdata <- ddply(df, c("Family", "Treatment","Type"), summarise,
                 N    = length(Abundance),
                 mean = mean(Abundance),
                 sd   = sd(Abundance),
                 se   = sd / sqrt(N)
)
```

```{r TreatmentPlot}
cdataplot=ggplot(Trtdata, aes(x=Treatment,y=mean))+geom_bar(aes(fill = Family),colour="black", stat="identity")+xlab("Treatment")+ylab("Relative Abundance (> 2%)")+theme(axis.title.x=element_blank())+ theme(axis.text.x = element_text(angle = 45, hjust = 1))+facet_wrap(~Type)+ scale_fill_manual(values=cbPalette)
cdataplot#+  scale_x_discrete(labels=c("0 hrs", "24 hrs", "48 hrs","72 hrs"))

#position = "fill"#+geom_errorbar(aes(ymin=mean-se,ymax=mean+se),color="black",width=1)+geom_line(size=1.5,linetype="dashed")+geom_point(size=6)+ylab(

#p <- ggbarplot(df, x = "Treatment", y = "Abundance",add = c("mean_se"),#"mean_se"
#               color = "black", palette = "cbPalette", facet.by="Family",
#               line.color = "gray", line.size = 0.4, short.panel.labs = TRUE, p.adjust.method = "fdr", fill= "Treatment") + stat_compare_means(aes(group = Treatment), label = "..p.signif..",label.y = 7) 

#p+ theme(axis.text.x = element_text(angle = 45, hjust = 1))+ylab("Relative abundance (> 1%)")+ theme(legend.position="none")
```


```{r}
Trtdata <- ddply(df, c("Family", "Treatment"), summarise,
                 N    = length(Abundance),
                 mean = mean(Abundance),
                 sd   = sd(Abundance),
                 se   = sd / sqrt(N)
)
cdataplot=ggplot(Trtdata, aes(x=Treatment,y=mean))+geom_bar(aes(fill = Treatment),colour="black", stat="identity")+ facet_grid(~Sample_Type)+xlab("Decomp Stage")+ylab("Relative Abundance (> 2%)") + theme(axis.text.x = element_text(angle = 45, hjust = 1))+theme(axis.title.x=element_blank())+facet_wrap(~Family)+geom_errorbar(aes(ymin=mean-se,ymax=mean+se))+scale_colour_manual(values=cbPalette)+scale_fill_manual(values=cbPalette)
#scale_x_discrete(labels=c("0 hrs", "24 hrs", "48 hrs","72 hrs"))+ theme(axis.text.x = element_text(angle = 45, hjust = 1))#+ scale_fill_manual(values=cbPalette)
cdataplot # position = "fill"#+geom_errorbar(aes(ymin=mean-se,ymax=mean+se),color="black",width=1)+geom_line(size=1.5,linetype="dashed")+geom_point(size=6)
theme_set(theme_bw(base_size = 18)+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())) 
```


```{r}
Means=compare_means(Abundance ~ Treatment, data = df, 
                    group.by = "Family", p.adjust.method = "fdr")
#head(Means)
keeps <- c("Family","group1","group2","p.adj","method","p.signif")
keeps=Means[keeps]
#keeps


test3 <- list('Family'= keeps$Family,'group1'=keeps$group1,'group2'= keeps$group2,'p.adj'=keeps$p.adj,p.signif=keeps$p.signif,'Method'=keeps$method)
test3= as.data.frame(test3)
#test3
FilteredResults<-test3[!(test3$p.adj>0.1),]            
FilteredResults
```

##Genus level

```{r Summarizing3,warning=FALSE}
df <- psmelt(GenusLevel)
df$Abundance=df$Abundance*100
Trtdata <- ddply(df, c("Genus", "Treatment","Type"), summarise,
                 N    = length(Abundance),
                 mean = mean(Abundance),
                 sd   = sd(Abundance),
                 se   = sd / sqrt(N)
)
```

```{r TreatmentPlot4}
cdataplot=ggplot(Trtdata, aes(x=Treatment,y=mean))+geom_bar(aes(fill = Genus),colour="black", stat="identity")+xlab("Treatment")+ylab("Relative Abundance (> 2%)") + theme(axis.text.x = element_text(angle = 0, hjust = 0.5))+theme(axis.title.x=element_blank())+ theme(axis.text.x = element_text(angle = 45, hjust = 1))+facet_wrap(~Type)+ scale_fill_manual(values=cbPalette)
cdataplot#+  scale_x_discrete(labels=c("0 hrs", "24 hrs", "48 hrs","72 hrs"))
#position = "fill"#+geom_errorbar(aes(ymin=mean-se,ymax=mean+se),color="black",width=1)+geom_line(size=1.5,linetype="dashed")+geom_point(size=6)+ylab(

#p <- ggbarplot(df, x = "Treatment", y = "Abundance",add = c("mean_se"),#"mean_se"
#               color = "black", palette = "cbPalette", facet.by="Family",
#               line.color = "gray", line.size = 0.4, short.panel.labs = TRUE, p.adjust.method = "fdr", fill= "Treatment") + stat_compare_means(aes(group = Treatment), label = "..p.signif..",label.y = 7) 

#p+ theme(axis.text.x = element_text(angle = 45, hjust = 1))+ylab("Relative abundance (> 1%)")+ theme(legend.position="none")
Trtdata <- ddply(df, c("Genus", "Treatment"), summarise,
                 N    = length(Abundance),
                 mean = mean(Abundance),
                 sd   = sd(Abundance),
                 se   = sd / sqrt(N)
)
cdataplot=ggplot(Trtdata, aes(x=Treatment,y=mean))+geom_bar(aes(fill = Treatment),colour="black", stat="identity")+ facet_grid(~Genus)+xlab("Decomp Stage")+ylab("Relative Abundance (> 2%)") + theme(axis.text.x = element_text(angle = 45, hjust = 1))+theme(axis.title.x=element_blank())+facet_wrap(~Genus)+geom_errorbar(aes(ymin=mean-se,ymax=mean+se))+scale_colour_manual(values=cbPalette)+scale_fill_manual(values=cbPalette)
#scale_x_discrete(labels=c("0 hrs", "24 hrs", "48 hrs","72 hrs"))+ theme(axis.text.x = element_text(angle = 45, hjust = 1))+ scale_fill_manual(values=cbPalette)
cdataplot
Means=compare_means(Abundance ~ Treatment, data = df, 
                    group.by = "Genus", p.adjust.method = "fdr")
#head(Means)
keeps <- c("Genus","group1","group2","p.adj","method","p.signif")
keeps=Means[keeps]
#keeps


test3 <- list('Genus'= keeps$Genus,'group1'=keeps$group1,'group2'= keeps$group2,'p.adj'=keeps$p.adj,p.signif=keeps$p.signif,'Method'=keeps$method)
test3= as.data.frame(test3)
#test3
FilteredResults<-test3[!(test3$p.adj>0.1),]            
FilteredResults
```


#Ordinations
-All ordinations here were conducted using weighted unifrac (for other distince metrics see QIIME outputs)

-Ellipses represent 95% CI for the mean of each group

##PCoA 
###By Treatment
```{r PCoA, warning=FALSE}
ord=ordinate(physeq,"PCoA", "wunifrac")
ordplot=plot_ordination(physeq, ord,"samples", color="Treatment")+geom_point(size=4)+scale_colour_manual(values=cbPalette)+scale_fill_manual(values=cbPalette)
ordplot+ stat_ellipse(type= "norm",geom = "polygon", alpha = 1/4, aes(fill = Treatment))+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())#+ theme(legend.justification=c(1,0), legend.position=c(1,0))

```

```{r PCoA2, warning=FALSE}
ord=ordinate(physeq,"PCoA", "wunifrac")
ordplot=plot_ordination(physeq, ord,"samples", color="Treatment")+geom_point(size=4)+scale_colour_manual(values=cbPalette)+scale_fill_manual(values=cbPalette)
ordplot+ stat_ellipse(type= "norm",geom = "polygon", alpha = 1/4, aes(fill = Treatment))+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+facet_wrap(~Type)#+ theme(legend.justification=c(1,0), legend.position=c(1,0))

```



#PERMANOVAs

##Homogenieity of Multivariate Dispersions
```{r, warning=FALSE}
GPdist=phyloseq::distance(physeq, "wunifrac")
beta=betadisper(GPdist, sample_data(physeq)$Treatment)
permutest(beta)
boxplot(beta)
```

##Treatment x Location
``` {r PERMANOVAs,warning=FALSE}
GPdist=phyloseq::distance(physeq, "wunifrac")
#MONMDS= ordinate(physeq, "NMDS",GPdist)
```

```{r}
adonis(GPdist ~ Treatment*Type, as(sample_data(physeq), "data.frame"))
```
